Checking the CIDnetworks Suite: Stochastic Block Model (SBM)
========================================================

A.C. Thomas, November 21

`r library(CIDnetworks)`

Stochastic Block Model: Component Alone
---------------------------------------

$Y_{ij} = S_i B S_j$

Component alone. Specify a model for simulation, create the data, and run the Gibbs sampler.

```{r}
test.sbm <- SBM(2, n.nodes=30, b.vector=c(1,-1,1), generate=TRUE, residual.variance=0.0001, restrict.and.shift=FALSE)
test.sbm$plot()
test.sbm.pieces <- test.sbm$pieces()
test.sbm.value <- test.sbm$value()

test.sbm.gibbs <- test.sbm$gibbs.full(draws=500, report=100, burnin=100, thin=3, make.random.start=TRUE)
test.sbm$gibbs.plot(test.sbm.gibbs)
```

Now, process the output in a way that will be meaningful to testing the model.

```{r}
test.sbm.gibbs.unlist <- list.output.to.matrices (test.sbm.gibbs)
test.sbm.gibbs.value <- test.sbm$gibbs.value(test.sbm.gibbs)
```

Test each of the parameters. First, the log likelihood:

```{r fig.width=12}
plot(test.sbm.gibbs.unlist$log.lik, main="Log-likelihood")
```

How do the network values look?

```{r fig.width=12}
colrange <- range(c(test.sbm.value, apply(test.sbm.gibbs.value, 1, mean)))
netplot(test.sbm$edge.list, test.sbm.value, extremes=colrange, colvalues=20, main="Oracle values")
netplot(test.sbm$edge.list, apply(test.sbm.gibbs.value, 1, mean), extremes=colrange, colvalues=20, main="Fitted values")
plot(test.sbm.value, apply(test.sbm.gibbs.value, 1, mean))
```


Gaussian within CID:

```{r}
test.sbm.g <- CID.generate (50, components=list(SBM(2, b.vector=c(1,-1,1))), intercept=1, residual.variance=0.02, generate=TRUE, class="gaussian")
test.sbm.g$plot()
test.sbm.g.pieces <- test.sbm.g$pieces()
test.sbm.g.value <- test.sbm.g$value()
test.sbm.g.gibbs <- test.sbm.g$gibbs.full(draws=300, report=100, burnin=300, make.random.start=TRUE)
par(mfrow=c(2,2))
test.sbm.g$gibbs.plot(test.sbm.g.gibbs)
```

```{r fig.width=12}
test.sbm.g.gibbs.value <- test.sbm.g$gibbs.value(test.sbm.g.gibbs)

color.range <- range(c(test.sbm.g.value, apply(test.sbm.g.gibbs.value, 1, mean)))
netplot(test.sbm.g$edge.list, test.sbm.g.value, colvalues=20, main="Oracle values", extremes=color.range)
netplot(test.sbm.g$edge.list, apply(test.sbm.g.gibbs.value, 1, mean), colvalues=20, main="Fitted values", extremes=color.range)

plot(test.sbm.g.value, apply(test.sbm.g.gibbs.value, 1, mean))
```


Binary within CID:

```{r}
test.sbm.b <- CID.generate (100, components=list(SBM(2, b.vector=c(1,-1,1))), intercept=0, generate=TRUE)
test.sbm.b.pieces <- test.sbm.b$pieces()
test.sbm.b.value <- test.sbm.b$value()

test.sbm.b.gibbs <- test.sbm.b$gibbs.full(draws=2000, report=100, burnin=500, make.random.start=TRUE)
par(mfrow=c(2,2))
test.sbm.b$gibbs.plot(test.sbm.b.gibbs)
```

```{r fig.width=12}
test.sbm.b.gibbs.value <- test.sbm.b$gibbs.value(test.sbm.b.gibbs)

color.range <- range(c(test.sbm.b.value, apply(test.sbm.b.gibbs.value, 1, mean)))
netplot(test.sbm.b$edge.list, test.sbm.b.value, colvalues=20, main="Oracle values", extremes=color.range)
netplot(test.sbm.b$edge.list, apply(test.sbm.b.gibbs.value, 1, mean), colvalues=20, main="Fitted values", extremes=color.range)

plot(test.sbm.b.value, apply(test.sbm.b.gibbs.value, 1, mean))
```

